---
title: "HW 3R"
author: "Brendon Krupa"
date: "11/28/2019"
output: word_document
---

## The Data

```{r}
sig <- matrix(c(.1, .02, .01, .02, .11, .03, .01, .03, .2), ncol = 3, byrow = TRUE)*100
sig
inv_sig <- solve(sig)

p <- matrix(c(.045, .003, .0285))*100

e <- matrix(c(1, 1, 1))
e

A = as.numeric(t(p) %*% inv_sig %*% e)
A

B = as.numeric(t(p) %*% inv_sig %*% (p))
B

C = as.numeric(t(e) %*% inv_sig %*% e)
C

detM = as.numeric(B*C - A^2)
detM

g = (1/detM) * ((B * (inv_sig %*% e)) - A * (inv_sig %*% p))
g

h = ((1/detM) * (C * (inv_sig %*% p) - A * (inv_sig %*% e)))
h
```

## Markowitz Mean-Variance with Risky Securities

# Global Minimal Variance Portfolio
```{r}
Pstar_var = as.numeric(1/C)
Pstar_var

Pstar_ret = as.numeric(A/C)
Pstar_ret


wPstar = g + (h * Pstar_ret)
wPstar
```

Global Min Variance Portfolio with Optimizer

```{r}
library("quadprog")
b <- c(1, Pstar_ret)
Amat <- matrix(c(e, p), 3, 2)

sol <- solve.QP(sig, c(0,0,0), Amat, b, meq = 2)
sol$solution
```

# Portfolio 1
Target Return of p1/3

```{r}
P1 = p[1]/3
P1

wP1 = g + (h * P1)
wP1/sum(wP1)
```

Portfolio 1 with Optimizer

```{r}
b1 <- c(1, P1)

sol1 <- solve.QP(sig, c(0,0,0), Amat, b1, meq = 2)
wP1 <- sol1$solution
wP1
```

# Portfolio 2
Target Return of p2 + p3

```{r}
P2 = p[2] + p[3]
P2

wP2 = g + (h * P2)
wP2/sum(wP2)
```

Portfolio 2 with Optimizer

```{r}
b2 <- c(1, P2)

sol2 <- solve.QP(sig, c(0,0,0), Amat, b2, meq = 2)
wP2 <- sol2$solution
wP2
```

# Portfolio 3

```{r}
P3 <- (p[1] + p[2] + p[3])/3 
alpha <- (P3 - P2)/(P1 - P2)
alpha

wP3 <- (alpha * wP1) + (1 - alpha)*wP2
wP3
```

Portfolio 3 with Optimizer

```{r}
b3 <- c(1, P3)

sol3 <- solve.QP(sig, c(0,0,0), Amat, b3, meq = 2)
wP3 <- sol3$solution
wP3
```

# Mean-Variance Efficient Frontier

```{r}
plot(sqrt(t(wPstar) %*% sig %*% wPstar), t(p) %*% wPstar, xlab = 'Sigma (%)', ylab = 'Return (%)', main = 'Efficient Frontier', col = 'green', type = 'p', xlim = c(.2, 5), ylim = c(.4, 10))
lines(sqrt(t(wP1) %*% sig %*% wP1), t(p) %*% wP1, type = 'p', col = 'gray43')
lines(sqrt(t(wP2) %*% sig %*% wP2), t(p) %*% wP2, type = 'p', col = 'maroon1')
lines(sqrt(t(wP3) %*% sig %*% wP3), t(p) %*% wP3, type = 'p', col = 'darkorange3')

points(sqrt(diag(sig)), p, col = 'lightskyblue')
text(sqrt(diag(sig)), c(p[1] - .4, p[2] + .4,  p[3] + .4), col = 'lightskyblue', labels = c('S1', 'S2', 'S3'))
rets <- c()
sigs <- c()
for(t in seq(0, 10, length.out = 1000)){
  alpha <- (t - P2)/(P1 - P2)
  wP <- (alpha * wP1) + (1 - alpha)*wP2
  sigma <- sqrt(t(wP) %*% sig %*% wP)
  rets <- c(rets, t)
  sigs <- c(sigs, sigma)
}
lines(sigs, rets, type = 'l', col = 'navyblue')
abline(b = sqrt(detM/C), a = A/C, col = 'salmon')
abline(b = -sqrt(detM/C), a = A/C, col = 'salmon')
legend(.5, 10, legend = c("Min Variance", "Portfolio 1", "Portfolio 2", "Portfolio 3", "Efficient Frontier", "Asymptotes"), col = c("green", "gray43", "maroon1", "darkorange3", "navyblue", "salmon"), lty=1:2, cex=0.8)
```


## Markowitz Mean-Variance with Risky and Risk-Free Securities

```{r}
P0 <- 0.6
PT <- A/C - (detM/(C^2))/(P0-A/C)
pxs <- p - P0
H <- t(pxs) %*% inv_sig %*% pxs
sigT <- -sqrt(H)/(C * (P0 - A/C))

PT
sigT
```

Tangent Portfolio Weights

```{r}
pxs <- p - P0*e
wT <- (1/(A - C*P0))*(inv_sig %*% pxs)
wT

t(p) %*% wT
sqrt(t(wT) %*% sig %*% wT)
```

# Efficient Frontier

```{r}
plot(sigs, rets, type = 'l', col = 'navyblue', xlab = 'Sigma (%)', ylab = 'Return (%)', main = 'Efficient Frontier', xlim = c(.2, 5), ylim = c(.4, 10))

points(c(sqrt(t(wP1) %*% sig %*% wP1), sqrt(t(wP2) %*% sig %*% wP2), sqrt(t(wP3) %*% sig %*% wP3), sigT), c(t(p) %*% wP1, t(p) %*% wP2, t(p) %*% wP3, PT), col = 'green')
text(c(sqrt(t(wP1) %*% sig %*% wP1) +.2, sqrt(t(wP2) %*% sig %*% wP2) + .2, sqrt(t(wP3) %*% sig %*% wP3), sigT), c(t(p) %*% wP1, t(p) %*% wP2, t(p) %*% wP3 - .5, PT - .4), col = 'green', labels = c('P1', 'P2', 'P3', 'PT'))

points(sqrt(t(wPstar) %*% sig %*% wPstar), t(p) %*% wPstar, col = 'darkgoldenrod3')
text(sqrt(t(wPstar) %*% sig %*% wPstar) - .3, t(p) %*% wPstar, col = 'darkgoldenrod3', labels = c('Min Var'))
abline(b = sqrt(detM/C), a = A/C, col = 'salmon')
abline(b = -sqrt(detM/C), a = A/C, col = 'salmon')
points(sqrt(diag(sig)), p, col = 'lightskyblue')
text(sqrt(diag(sig)), c(p[1] - .4, p[2] + .4,  p[3] + .4), col = 'lightskyblue', labels = c('S1', 'S2', 'S3'))

abline(b = sqrt(H), a = P0, col = 'mediumpurple1')
abline(b = -sqrt(H), a = P0, col = 'mediumpurple1')
legend(1, 10, legend = c("Efficient Frontier", "Asymptotes", "Risk-Free EF"), col = c("navyblue", "salmon", "mediumpurple1"), lty=1:2, cex=0.8)
```

The efficient frontiers intersect at a single point, the tangent portfolio. This tangent portfolio represents the market portfolio that mimics the behavior of the market while only being invested in four securities. This second efficient frontier is the capital market line and therefore must intersect the risky efficient frontier at a single point that represents the market portfolio.

# Finding P4

```{r}
P4 <- 5
alpha4 <- (P4 - PT)/(P0 - PT)
sig4 <- (1 - alpha4)*sigT
alpha4
sig4
```

Finding Weights

```{r}
w4 <- c(alpha4, (1-alpha4)*wT)
t(c(P0,p)) %*% w4
w4
```


# Finding P5

```{r}
sig5 <- sqrt(16)
alpha5 <- as.numeric(1 - sig5/sigT)
P5 <- alpha5*P0 + (1 - alpha5)*PT
alpha5
P5
```

Finding Weights

```{r}
w5 <- c(alpha5, (1-alpha5)*wT)
t(c(P0,p)) %*% w5
w5
```

# Adding to Graph 

```{r}
plot(sigs, rets, type = 'l', col = 'navyblue', xlab = 'Sigma (%)', ylab = 'Return (%)', main = 'Efficient Frontier', xlim = c(.2, 5), ylim = c(.4, 10))

points(c(sqrt(t(wP1) %*% sig %*% wP1), sqrt(t(wP2) %*% sig %*% wP2), sqrt(t(wP3) %*% sig %*% wP3), sigT, sig4, sig5), c(t(p) %*% wP1, t(p) %*% wP2, t(p) %*% wP3, PT, P4, P5), col = 'green')
text(c(sqrt(t(wP1) %*% sig %*% wP1) +.2, sqrt(t(wP2) %*% sig %*% wP2) + .2, sqrt(t(wP3) %*% sig %*% wP3), sigT, sig4, sig5), c(t(p) %*% wP1, t(p) %*% wP2, t(p) %*% wP3 - .5, PT - .4, P4 + .4, P5 + .4), col = 'green', labels = c('P1', 'P2', 'P3', 'PT', 'P4', 'P5'))

points(sqrt(t(wPstar) %*% sig %*% wPstar), t(p) %*% wPstar, col = 'darkgoldenrod3')
text(sqrt(t(wPstar) %*% sig %*% wPstar) - .3, t(p) %*% wPstar, col = 'darkgoldenrod3', labels = c('Min Var'))
abline(b = sqrt(detM/C), a = A/C, col = 'salmon')
abline(b = -sqrt(detM/C), a = A/C, col = 'salmon')
points(sqrt(diag(sig)), p, col = 'lightskyblue')
text(sqrt(diag(sig)), c(p[1] - .4, p[2] + .4,  p[3] + .4), col = 'lightskyblue', labels = c('S1', 'S2', 'S3'))

abline(b = sqrt(H), a = P0, col = 'mediumpurple1')
abline(b = -sqrt(H), a = P0, col = 'mediumpurple1')
legend(1, 10, legend = c("Efficient Frontier", "Asymptotes", "Risk-Free EF"), col = c("navyblue", "salmon", "mediumpurple1"), lty=1:2, cex=0.8)
```

